
---
title: "Interface with other single-cell analysis toolkits"
author: "Suoqin Jin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    theme: united
mainfont: Arial
vignette: >
  %\VignetteIndexEntry{Interface with other single-cell analysis toolkits}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Converting between single-cell objects (Seurat, SingleCellExperiment and anndata objects)
There are several excellent tools to convert between [Seurat object](https://satijalab.org/seurat/), [SingleCellExperiment object](https://www.bioconductor.org/packages/release/bioc/vignettes/SingleCellExperiment/inst/doc/intro.html), and [anndata object](https://github.com/dynverse/anndata). Here are two examples: one is [Seurat R package](https://satijalab.org/seurat/v3.1/conversion_vignette.html) from Satija Lab,  another is [zellkonverter R package](https://github.com/theislab/zellkonverter) from Theis Lab, and [sceasy](https://github.com/cellgeni/sceasy). 

Below we shows how to extract the CellChat input files as data matrix from other existing single-cell analysis toolkits, including [Seurat](https://satijalab.org/seurat/) and [Scanpy](https://icb-scanpy.readthedocs-hosted.com/en/stable/tutorials.html). 


# TROUBLESHOOTING when starting from a AnnData object
If the RStudio encounters FATAL ERROR when creating a CellChat object from a AnnData/Scanpy object, users can save the required data files `data.input` and `meta` in user's local computer and then reload them for CellChat analysis. 

If the RStudio encounters FATAL ERROR when reading .h5ad file using zellkonverter::readH5AD, users can first insall the anndata R package via `install.packages("anndata")`. 

# Part I: Prepare required input data matrix and meta data for creating a CellChat object
## Required data inputs
CellChat requires two user inputs: one is the gene expression data of cells, and the other is the user assigned cell labels.

- *For the gene expression data matrix*, genes should be in rows with rownames and cells in columns with colnames. Normalized data is required as input for CellChat analysis, e.g., library-size normalization and then log-transformed with a pseudocount of 1. If user provides count data, we provide a `normalizeData` function to account for library size. 

- *For the cell group information*, a dataframe with rownames is required as input for CellChat. 

In addition to taking a count data matrix as an input, we also provide instructions for how to prepare CellChat input files from other existing single-cell analysis toolkits, including Seurat and Scanpy. Please start to prepare the input data by following option A when the normalized count data and meta data are available, option B when the Seurat object is available, option C when the SingleCellExperiment object is available, and option D when the Scanpy object is available.  


## (A) Starting from a count data matrix
```{r}
load("./tutorial/data_humanSkin_CellChat.rda")
data.input = data_humanSkin$data # normalized data matrix
meta = data_humanSkin$meta # a dataframe with rownames containing cell mata data
# Subset the input data for CelChat analysis
cell.use = rownames(meta)[meta$condition == "LS"] # extract the cell names from disease data
data.input = data.input[, cell.use]
meta = meta[cell.use, ]
```

## (B) Starting from a Seurat object
The normalized count data and cell group information can be obtained from the Seurat object by
```{r}
data.input <- GetAssayData(seurat_object, assay = "RNA", slot = "data") # normalized data matrix
labels <- Idents(seurat_object)
meta <- data.frame(group = labels, row.names = names(labels)) # create a dataframe of the cell labels
```

## (C) Starting from a SingleCellExperiment object
The normalized logcount data and cell group information can be obtained from the SingleCellExperimen object by
```{r}
data.input <- SingleCellExperiment::logcounts(object) # normalized data matrix
meta <- as.data.frame(SingleCellExperiment::colData(object)) # extract a dataframe of the cell labels
meta$labels <- meta[["sce.clusters"]]
```

Below is a reproducible example: 
```{r}
library(SingleCellExperiment)
library(basilisk)
library(scRNAseq)
sce <- SegerstolpePancreasData()
sce
counts <- assay(sce, "counts")
libsizes <- colSums(counts)
size.factors <- libsizes/mean(libsizes)
logcounts(sce) <- log2(t(t(counts)/size.factors) + 1)

data.input <- SingleCellExperiment::logcounts(sce) # normalized data matrix
meta <- as.data.frame(SingleCellExperiment::colData(sce)) # extract a dataframe of the cell labels
meta$labels <- meta[["cell.type"]]
```

## (D) Starting from a AnnData object
anndata provides a python class that can be used to store single-cell data, which has been widely used in various Python tools such as the scanpy python package. 

### Using anndata R package
[anndata for R](https://github.com/dynverse/anndata), which is a reticulate wrapper for the anndata Python package, was recently published in CRAN, which makes it much convenient to read from and write to the h5ad file format in R. We show how to extract the required data inputs for CellChat analysis using anndata R package. 
```{r}
# read the data into R using anndata R package
library(anndata)
ad <- read_h5ad("scanpy_object.h5ad")
# access count data matrix
counts <- t(as.matrix(ad$X))
# normalize the count data if the normalized data is not available in the .h5ad file
library.size <- Matrix::colSums(counts)
data.input <- as(log1p(Matrix::t(Matrix::t(counts)/library.size) * 10000), "dgCMatrix")

# access meta data
meta <- ad$obs
# save the data for CellChat analysis
save(data.input, meta, file = "test_data.RData")
```

### Using reticulate R package
User can also read the data into R using the reticulate package to import the anndata module.
```{r}
# read the data into R using the reticulate package to import the anndata module
library(reticulate)
ad <- import("anndata", convert = FALSE)
ad_object <- ad$read_h5ad("scanpy_object.h5ad")
# access normalized data matrix
data.input <- t(py_to_r(ad_object$X))
rownames(data.input) <- rownames(py_to_r(ad_object$var))
colnames(data.input) <- rownames(py_to_r(ad_object$obs))
# access meta data
meta.data <- py_to_r(ad_object$obs)
meta <- meta.data

```
Please also check the suggestions by other USERS (https://github.com/sqjin/CellChat/issues/300#issuecomment-1116343516)

### Exporting required data directly from Python
User can also export expression matrix and meta data from h5ad in Python.
```{r}
pd.DataFrame(ad.var.index).to_csv(os.path.join(destination, "genes.tsv" ),   sep = "\t", index_col = False)
pd.DataFrame(ad.obs.index).to_csv(os.path.join(destination, "barcodes.tsv"), sep = "\t", index_col = False)
ad.obs.to_csv(os.path.join(destination, "metadata.tsv"), sep = "\t", index_col = True)
adata.T.to_df().to_csv('matrix.csv')
```

Please see more discussion on how to export expression matrix and meta data from h5ad (https://github.com/scverse/scanpy/issues/262)

# Part II: Create a CellChat object directly from Seurat, SingleCellExperiment or Scanpy object
From CellChat version 0.5.0, **USERS can create a new CellChat object from Seurat or SingleCellExperiment object.** . If input is a Seurat or SingleCellExperiment object, the meta data in the object will be used by default and USER must provide `group.by` to define the cell groups. e.g, group.by = "ident" for the default cell identities in Seurat object. 

Please check the examples in the documentation of `createCellChat` for details via `help(createCellChat)`. 

**NB: If USERS load previously calculated CellChat object (version < 0.5.0),  please update the object via `updateCellChat`** 

Create a CellChat object by following option A when taking the digital gene expression matrix and cell label information as input, option B when taking a Seurat object as input, option C when taking a SingleCellExperiment object as input, and option D when taking a AnnData object as input. See details in the subsection of Required input data. User can initialize a CellChat object using the createCellChat function as follows:

## (A) Starting from the digital gene expression matrix and cell label information
Upon extracting the required CellChat input files, then create a CellChat object and start the analysis.

```{r}
library(CellChat)
cellchat <- createCellChat(object = data.input, meta = meta, group.by = "labels")

```
If cell mata information is not added when creating CellChat object, Users can also add it later using `addMeta`, and set the default cell identities using `setIdent`.
```{r eval = FALSE}
cellchat <- addMeta(cellchat, meta = meta, meta.name = "labels")
cellchat <- setIdent(cellchat, ident.use = "labels") # set "labels" as default cell identity
levels(cellchat@idents) # show factor levels of the cell labels
```

## (B) Starting from a Seurat object
```{r}
library(CellChat)
cellChat <- createCellChat(object = seurat.obj, group.by = "ident", assay = "RNA")

```

## (C) Starting from a SingleCellExperiment object
```{r}
library(CellChat)
cellChat <- createCellChat(object = sce.obj, group.by = "sce.clusters")

```

## (D) Starting from a AnnData object
```{r}
sce <- zellkonverter::readH5AD(file = "adata.h5ad")
sce
# retrieve all the available assays within sce object
assayNames(sce)
# added a new assay entry "logcounts" if not available
counts <- assay(sce, "X") # make sure this is the original count data matrix
library.size <- Matrix::colSums(counts)
logcounts(sce) <- log1p(Matrix::t(Matrix::t(counts)/library.size) * 10000)

# extract a cell meta data
meta <- as.data.frame(SingleCellExperiment::colData(sce)) #
# save the sce object for CellChat analysis
save(sce, file = "sce_object.RData")

# create a CellChat object from the sce object
cellChat <- createCellChat(object = sce.obj, group.by = "clusters.final")
```


# Part III: Interface with other cell-cell communication analysis toolkits
## Use other ligand-receptor interaction databases for CellChat analysis

To use other ligand-receptor interaction databases for CellChat analysis, users need to convert the source database to the format of CellChatDB using the function named `updateCellChatDB`. Please check the tutorial named "Update-CellChatDB.html" (https://htmlpreview.github.io/?https://github.com/jinworks/CellChat/blob/master/tutorial/Update-CellChatDB.html) for more details on updating CellChatDB by integrating new ligand-receptor pairs from other resources or utilizing a custom ligand-receptor interaction database. 

## Use other computed cell-cell interaction scores for CellChat analysis

To use CellChat for downstream analysis of cell-cell communication from other cell-cell communication tools, users can use the function named `updateCCC_score`, which allows users to provide customized cell-cell interaction scores. Please check the detailed documentation of this function via `help(updateCCC_score)` . 
