
---
title: "Update CellChatDB by integrating new ligand-receptor pairs from other resources or utilizing a custom ligand-receptor interaction database"
author: "Suoqin Jin"
date: "01 November, 2023"
output:
  html_document:
    toc: true
    theme: united
mainfont: Arial
vignette: >
  %\VignetteIndexEntry{Update CellChatDB by integrating new ligand-receptor pairs from other resources or utilizing a custom ligand-receptor interaction database}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  root.dir = './'
)
knitr::opts_chunk$set(eval = FALSE)
```

# Part I: update CellChatDB using the function `updateCellChatDB`

CellChat v2 provides the function `updateCellChatDB` to allow users to update CellChatDB by integrating new ligand-receptor pairs from other resources or utilizing a custom ligand-receptor interaction database. This function can be used as follows: 
```{r eval=FALSE}
updateCellChatDB(
  db,
  gene_info = NULL,
  other_info = NULL,
  merged = FALSE,
  species_target = NULL
)
```
The minimum input of this function is a data frame with two columns named as `ligand` and `receptor`, and other inputs include classified pathway, complex, cofactor and gene information. Users can check the documentation of this function for more details of the required input data by typing `help(updateCellChatDB)` in RStudio.

Here we describe how to integrate other resources by taking another database CellTalkDB as an example. After downloading the CellTalkDB from https://github.com/ZJUFanLab/CellTalkDB, follow the four steps as follows. 

## Load the required libraries
```{r message=FALSE,warning=FALSE}
library(CellChat)
options(stringsAsFactors = FALSE)
```

## Step 1: Load the customized ligand-receptor pairs and the gene information
The first input `db` is a data frame with as least two columns named as `ligand` and `receptor`. We highly suggest users to provide a column of pathway information named 'pathway_name' associated with each L-R pair. Other optional columns include 'interaction_name' and 'interaction_name_2'. The default columns of CellChatDB can be checked via 'colnames(CellChatDB.human$interaction)'.

The second input `gene_info` is a data frame with at least one column named as 'Symbol'. "When setting gene_info = NULL, the input 'species_target' should be provided: either 'human' or 'mouse'. 

The third input `other_info` is a list consisting of other information including a dataframe named as 'complex' and a dataframe named as 'cofactor'. This additional information is not necessary. If other_info is provided, the 'complex' and 'cofactor' are dataframes with defined rownames. 

```{r eval = FALSE}
# Load the dataframe consisting of customized ligand-receptor pairs
db.user <- readRDS("./CellTalkDB-master/database/human_lr_pair.rds")
# Load the dataframe consisting of gene information (optional)
gene_info <- readRDS("./CellTalkDB-master/data/human_gene_info.rds")
```

## Step 2: Formulate the input files to be compatible with CellChatDB
Please check above for the basic requirement of the input files. In addition, users can check the detailed database information in CellChatDB. 

* The main ligand-receptor interaction information is stored in `CellChatDB$interaction`. Columns of this dataframe includes information of ligands, receptors and co-factors. 

* ` CellChatDB$geneInfo` contains all gene information in mouse or other species and it should have a column named ‘Symbol’, which is not necessary to be changed when updating CellChatDB. 

* Optional information is stored in `CellChatDB$complex` and `CellChatDB$cofactor`, which include any ligand complex, receptor complex and cofactors. Users need to make sure that user-defined complex/cofactor names are the same in `CellChatDB$interaction`. 

```{r}
# Modify the colnames because of incompatible with `colnames(CellChatDB.human$interaction)`
colnames(db.user) <- plyr::mapvalues(colnames(db.user), from = c("ligand_gene_symbol","receptor_gene_symbol","lr_pair"), to = c("ligand","receptor","interaction_name"), warn_missing = TRUE)
```

## Step 3: Update CellChatDB by running the function `updateCellChatDB`
```{r}
# Create a new database by typing one of the following commands
# Use user-provided gene information
db.new <- updateCellChatDB(db = db.user, gene_info = gene_info)
# Use built-in gene information of CellChatDB
db.new <- updateCellChatDB(db = db.user, gene_info = NULL, species_target = "human")

# Alternatively, users can integrate the customized L-R pairs into the built-in CellChatDB 
db.new <- updateCellChatDB(db = db.user, merged = TRUE, species_target = "human")

```
## Step 4: Use the new database in CellChat analysis or re-build CellChat package
Re-build CellChat package by updating the database as follows
```{r eval=FALSE}
# Users can now use this new database in CellChat analysis 
cellchat@DB <- db.new

# Users can save the new database for future use
save(db.new, file = "CellChatDB.human_user.rda")

# Users can also re-build CellChatDB in CellChat package
setwd("/Users/$USERS/Downloads/CellChat-master") # This is the folder of CellChat package downloaded from Github
CellChatDB.mouse <- CellChatDB
usethis::use_data(CellChatDB.mouse, overwrite = TRUE)

# If working on a human dataset, do following:
# CellChatDB.human <- CellChatDB
# usethis::use_data(CellChatDB.human, overwrite = TRUE)
```

# Part II: update CellChatDB by manually modifying the required files
Here we outline the steps to update CellChatDB by manually adding user-defined ligand-receptor pairs. To do so, the format of the users’ lists must be compatible with the input files of CellChatDB. 

## Load the required libraries
```{r message=FALSE,warning=FALSE}
library(CellChat)
options(stringsAsFactors = FALSE)
```

## Step 1: Access the ligand-receptor interaction information in CellChatDB
Extract the database information in CellChatDB and then save them in a local computer, including four files: 'geneInfo.csv', 'interaction_input_CellChat.csv', 'complex_input_CellChat.csv', 'and cofactor_input_CellChat.csv'. Users can do it by running the following codes in Rstudio:

```{r}
CellChatDB <- CellChatDB.mouse # set CellChatDB <- CellChatDB.human if working on the human dataset
interaction_input <- CellChatDB$interaction
complex_input <- CellChatDB$complex
cofactor_input <- CellChatDB$cofactor
geneInfo <- CellChatDB$geneInfo
write.csv(interaction_input, file = "interaction_input_CellChatDB.csv")
write.csv(complex_input, file = "complex_input_CellChatDB.csv")
write.csv(cofactor_input, file = "cofactor_input_CellChatDB.csv")
write.csv(geneInfo, file = "geneInfo_input_CellChatDB.csv")

```

## Step 2: Update the required files by adding users’ curated ligand-receptor pairs
Update the above four .csv files by adding users’ curated ligand-receptor pairs. 

* The main file is 'interaction_input_CellChatDB.csv'. Users can first update the ligands, receptors and co-factors in the corresponding columns in 'interaction_input_CellChatDB.csv'. 

* Users can then update "complex_input_CellChatDB.csv" and "cofactor_input_CellChatDB.csv" if any ligand complex, receptor complex and cofactors are updated. Users need to make sure that user-defined complex/cofactor names are the same in 'interaction_input_CellChatDB.csv' and "complex_input_CellChatDB.csv", 'interaction_input_CellChatDB.csv' and " cofactor_input_CellChatDB.csv".

* "geneInfo_input_CellChatDB.csv" contains all gene information in mouse and it should have a column named ‘Symbol’, which does not need to be changed when updating CellChatDB. 


## Step 3: Update CellChatDB
Update CellChatDB once updating the four .csv files. Users can do it by running the following codes in Rstudio:
```{r}
options(stringsAsFactors = FALSE)
interaction_input <- read.csv(file = 'interaction_input_CellChatDB.csv', row.names = 1)
complex_input <- read.csv(file = 'complex_input_CellChatDB.csv', row.names = 1)
cofactor_input <- read.csv(file = 'cofactor_input_CellChatDB.csv', row.names = 1)
geneInfo <- read.csv(file = ' geneInfo_input_CellChatDB.csv', row.names = 1)
CellChatDB <- list()
CellChatDB$interaction <- interaction_input
CellChatDB$complex <- complex_input
CellChatDB$cofactor <- cofactor_input
CellChatDB$geneInfo <- geneInfo

```

## Step 4: Re-build CellChat package (optional)
Re-build CellChat package by updating the database as follows
```{r}
setwd("/Users/$USERS/Downloads/CellChat-master") # This is the folder of CellChat package downloaded from Github
CellChatDB.mouse <- CellChatDB
usethis::use_data(CellChatDB.mouse, overwrite = TRUE)

# If working on a human dataset, do following:
# CellChatDB.human <- CellChatDB
# usethis::use_data(CellChatDB.human, overwrite = TRUE)
```
